Dependent Object Types

### 2012-09-04 doc/candidacy/report.tex
a bit more about tribe paths

In~\cite{tribe}, Clarke et al. present a calculus for virtual
classes. Virtual classes allow {\it families} of classes to be
inherited, rather than just single classes. Classes are lexically
nested inside other classes. When a class is inherited, its nested
inner classes are inherited along with their methods and fields. {\it
  Further binding} enables the nested classes to be further
extended. Path-dependent types arise naturally to distinguish objects
coming from different families. Yet, thanks to {\it family
  polymorphism}, code written for one family also works for extension
of that family. In Tribe, path types can depend simultaneously on both
classes and objects. Additionally, paths can use an \code{out} field
to move from an object to the object which surrounds it. This
\code{out} feature enables ubiquitous access to an object's family
without the need to drag around family arguments.

@@

@@

In terms of theoretical results, there's a sketch of soundness in the
paper, but it's unclear whether it is correct. In particular, the
substitution lemma does not seem valid when null is the value to be
substituted and the expression has a type which contains the variable
to be substituted. In the follow-up paper~\cite{tribalo}, the
soundness proof sketch seems more plausible but then is tied to the
ownership system formalism. Decidability of subtyping, and hence
typechecking, has not yet been established.

>>>>

In~\cite{tribe}, Clarke et al. present a calculus for virtual
classes. Virtual classes allow {\it families} of classes to be
inherited, rather than just single classes. Classes are lexically
nested inside other classes. When a class is inherited, its nested
inner classes are inherited along with their methods and fields. {\it
  Further binding} enables the nested classes to be further
extended. Path-dependent types arise naturally to distinguish objects
coming from different families. Yet, thanks to {\it family
  polymorphism}, code written for one family also works for extensions
of that family. In Tribe, path types can depend simultaneously on both
classes and objects. Additionally, paths can use an \code{out} field
to move from an object to the object which surrounds it. This
\code{out} feature enables ubiquitous access to an object's family
without the need to drag around family arguments.

@@

In Tribe, types are a form of generalized paths: a final variable,
\code{this} or a class name then followed by a possibly empty sequence
of field or class selections. Given the code below, we can read the
type \code{kitt.Passenger.name} as the name of one of Kitt's
passengers (including the driver). Subtyping is natural:
\code{kitt.driver <: Car.driver <: Car.Passenger <: Car.Traveller <: Vehicle.Traveller}. Crucially, \code{kitt.driver} is not a subtype
of \code{karr.driver} and similarly for \code{kitt.Passenger} and
\code{karr.Passenger}.
\begin{lstlisting}
class Vehicle { class Traveller { ... } }
class Car extends Vehicle {
  class Passenger extends Traveller {
    class String { ... }
    final String name;
  }
  final Passenger driver;
}
final Car kitt;
final Car karr;
\end{lstlisting}

@@

In terms of theoretical results, there's a sketch of soundness in the
paper, but it's unclear whether it is correct. In particular, the
substitution lemma does not seem valid when \code{null} is the value to be
substituted and the expression has a type which contains the variable
to be substituted. In the follow-up paper~\cite{tribalo}, the
soundness proof sketch seems more plausible but then is tied to the
ownership system formalism. Decidability of subtyping, and hence
typechecking, has not yet been established.
